The Dynamics of Normalization Reweighting

### 2022-05-17 manuscript.Rmd
Included MEG analysis code

# reasonably compact code to check which packages are installed, install the missing ones, and activate all
packagelist <- c('knitr','remotes','tictoc') # list of CRAN packages
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
if (!'osfr' %in% installed.packages()[,1]){remotes::install_github("centerforopenscience/osfr")}
if (!'FourierStats' %in% installed.packages()[,1]){remotes::install_github("bakerdh/FourierStats")}
packagelist <- c(packagelist,'osfr','FourierStats')
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

@@

# check if a directory exists to store data files (ignored by git)
if (!dir.exists('temp/')){dir.create('temp/')}
if (!dir.exists('temp/VilidaiteProcessed/')){dir.create('temp/VilidaiteProcessed/')}

@@

>>>>

# reasonably compact code to check which packages are installed, install the missing ones, and activate all
packagelist <- c('knitr','remotes','tictoc','R.matlab') # list of CRAN packages
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
if (!'osfr' %in% installed.packages()[,1]){remotes::install_github("centerforopenscience/osfr")}
if (!'FourierStats' %in% installed.packages()[,1]){remotes::install_github("bakerdh/FourierStats")}
packagelist <- c(packagelist,'osfr','FourierStats')
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

@@

# check if directories exist to store data files (ignored by git)
if (!dir.exists('temp/')){dir.create('temp/')}
if (!dir.exists('temp/VilidaiteProcessed/')){dir.create('temp/VilidaiteProcessed/')}
if (!dir.exists('temp/raw/')){dir.create('temp/raw/')}
if (!dir.exists('temp/raw/temp/')){dir.create('temp/raw/temp/')}
if (!dir.exists('temp/MEGdata/')){dir.create('temp/MEGdata/')}

@@

samplerate <- 1001
targetF <- 5
maskF <- 7
    binwidth <- 1001
    tindex <- 1 + targetF*(binwidth/1001)
    mindex <- 1 + maskF*(binwidth/1001)

d <- dir(path = 'temp/MEGdata/', pattern = '*alltrials.mat', full.names=TRUE)
d2 <- dir(path = 'temp/MEGdata/', pattern = '*alltrials.mat', full.names=FALSE)

allresps <- array(0, dim=c(length(d),4,2,9))
allspectra <- array(0, dim=c(length(d),4,6006))
allwaves <- array(0, dim=c(length(d), 4, 9010))
MEGtimecourse7Hz <- array(0, dim=c(length(d),4,8009))
MEGtimecourse5Hz <- array(0, dim=c(length(d),4,8009))
allRnos <- NULL
for (s in 1:length(d)){
  allRnos[s] <- substr(d2[s],1,5)
  data <- readMat(d[s])

  for (cond in 1:4){
      for (t in 1:9){
        all1F <- NULL
        all2F <- NULL
        for (rep in 1:data$trialcounter[1,cond]){
          trial <- data$alltrials[cond,rep,]

        sample <- trial[(1+((t-1)*1001)):(t*1001)]
        fftsample <- fft(sample)/length(sample)

        all1F[rep] <- fftsample[targetF + 1]
        all2F[rep] <- fftsample[maskF + 1]
      }

        tempxy <- data.frame(Re(all1F),Im(all1F))
        D <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
        i <- which(D<3)
        allresps[s,cond,1,t] <- mean(all1F[i])

        tempxy <- data.frame(Re(all2F),Im(all2F))
        D <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
        i <- which(D<3)
        allresps[s,cond,2,t] <- mean(all2F[i])
    }

    meanwave <- colMeans(data$alltrials[cond,1:data$trialcounter[1,cond],])

    sample <- meanwave[1001:7006]
    allspectra[s,cond,] <- fft(sample)/length(sample)

    allwaves[s,cond,] <- meanwave

fspec1 <- NULL
fspec2 <- NULL
for (t in 1:(length(meanwave)-binwidth)){
  spec <- fft(meanwave[t:(t+binwidth-1)])/binwidth
  fspec1[t] <- spec[tindex]
  fspec2[t] <- spec[mindex]
}
    MEGtimecourse5Hz[s,cond,] <- fspec1
    MEGtimecourse7Hz[s,cond,] <- fspec2

  }
}

allresps <- abs(allresps)
meanspec <- apply(allresps,2:4,mean)

MEGmeantime5Hz <- matrix(0,nrow=4,ncol=8009) 
MEGmeantime7Hz <- matrix(0,nrow=4,ncol=8009) 
for (cond in 1:4){
MEGmeantime7Hz[cond,] <- apply(abs(MEGtimecourse7Hz[,cond,]),2,mean)
MEGmeantime5Hz[cond,] <- apply(abs(MEGtimecourse5Hz[,cond,]),2,mean)
}

qdata <- read.csv('~/Google Drive/Current work/Teaching/MSci project 2021/MSci project questionnaires 2021 (Responses) - Form Responses 1.csv')
qgroup <- as.factor(qdata[,5])
qRnos <- as.character(qdata[,2])
qAges <- qdata[,3]
qRnos[20] <- 'R5860'
qgroup[20] <- 'Yes'
qAges[20] <- 19

sortedgroups <- NULL
for (s in 1:length(allRnos)){sortedgroups[s] <- qgroup[which(qRnos==allRnos[s])]}
# sortedgroups now codes 1 = control, 2 = ASC in the order of the MEG data

lowAQ <- abs(apply(abs(allresps[which(sortedgroups==1),,,]),c(2,3,4),mean))
highAQ <- abs(apply(abs(allresps[which(sortedgroups==2),,,]),c(2,3,4),mean))

MEGmeantime5HzCON <- matrix(0,nrow=4,ncol=8009) 
MEGmeantime7HzCON <- matrix(0,nrow=4,ncol=8009) 
MEGmeantime5HzASD <- matrix(0,nrow=4,ncol=8009) 
MEGmeantime7HzASD <- matrix(0,nrow=4,ncol=8009) 
for (cond in 1:4){
MEGmeantime7HzCON[cond,] <- apply(abs(MEGtimecourse7Hz[which(sortedgroups==1),cond,]),2,mean)
MEGmeantime5HzCON[cond,] <- apply(abs(MEGtimecourse5Hz[which(sortedgroups==1),cond,]),2,mean)
MEGmeantime7HzASD[cond,] <- apply(abs(MEGtimecourse7Hz[which(sortedgroups==2),cond,]),2,mean)
MEGmeantime5HzASD[cond,] <- apply(abs(MEGtimecourse5Hz[which(sortedgroups==2),cond,]),2,mean)
}

# par(mfrow=c(2,3))

times <- ((1:8009)-500)/1000

plotlims <- c(-1,8,0,30)  # define the x and y limits of the plot (minx,maxx,miny,maxy)
ticklocsx <- -1:8    # locations of tick marks on x axis
ticklocsy <- seq(0,30,10)    # locations of tick marks on y axis
ticklabelsx <- -1:8        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2) 
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx) 
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (s)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

# lines(times,1000*meanspec[1,1,],lwd=2,col='red')
# lines(times,1000*meanspec[2,1,],lwd=2,col='green')
# lines(times,1000*meanspec[3,1,],lwd=2,col='blue')
# lines(times,1000*meanspec[4,1,],lwd=2,col='orange')

# lines(times,1000*MEGmeantime5Hz[1,],lwd=2,col='red')
# lines(times,1000*MEGmeantime5Hz[2,],lwd=2,col='green')
# lines(times,1000*MEGmeantime5Hz[3,],lwd=2,col='blue')
# lines(times,1000*MEGmeantime5Hz[4,],lwd=2,col='orange')
spl5a <- smooth.spline(times,1000*MEGmeantime5Hz[1,],df=20)
lines(spl5a$x,spl5a$y,col='red',lwd=2)
spl5b <- smooth.spline(times,1000*MEGmeantime5Hz[2,],df=20)
lines(spl5b$x,spl5b$y,col='green',lwd=2)
spl5c <- smooth.spline(times,1000*MEGmeantime5Hz[3,],df=20)
lines(spl5c$x,spl5c$y,col='blue',lwd=2)
spl5d <- smooth.spline(times,1000*MEGmeantime5Hz[4,],df=20)
lines(spl5d$x,spl5d$y,col='orange',lwd=2)

# lines(times,1000*MEGmeantime5HzASD[1,],lwd=2,col='red',lty=2)
# lines(times,1000*MEGmeantime5HzASD[2,],lwd=2,col='green',lty=2)
# lines(times,1000*MEGmeantime5HzASD[3,],lwd=2,col='blue',lty=2)
# lines(times,1000*MEGmeantime5HzASD[4,],lwd=2,col='orange',lty=2)
spl5aASD <- smooth.spline(times,1000*MEGmeantime5HzASD[1,],df=20)
lines(spl5aASD$x,spl5aASD$y,col='red',lwd=2,lty=2)
spl5bASD <- smooth.spline(times,1000*MEGmeantime5HzASD[2,],df=20)
lines(spl5bASD$x,spl5bASD$y,col='green',lwd=2,lty=2)
spl5cASD <- smooth.spline(times,1000*MEGmeantime5HzASD[3,],df=20)
lines(spl5cASD$x,spl5cASD$y,col='blue',lwd=2,lty=2)
spl5dASD <- smooth.spline(times,1000*MEGmeantime5HzASD[4,],df=20)
lines(spl5dASD$x,spl5dASD$y,col='orange',lwd=2,lty=2)

# lines(times,1000*MEGmeantime5HzCON[1,],lwd=2,col='red',lty=3)
# lines(times,1000*MEGmeantime5HzCON[2,],lwd=2,col='green',lty=3)
# lines(times,1000*MEGmeantime5HzCON[3,],lwd=2,col='blue',lty=3)
# lines(times,1000*MEGmeantime5HzCON[4,],lwd=2,col='orange',lty=3)
spl5aCON <- smooth.spline(times,1000*MEGmeantime5HzCON[1,],df=20)
lines(spl5aCON$x,spl5aCON$y,col='red',lwd=2,lty=3)
spl5bCON <- smooth.spline(times,1000*MEGmeantime5HzCON[2,],df=20)
lines(spl5bCON$x,spl5bCON$y,col='green',lwd=2,lty=3)
spl5cCON <- smooth.spline(times,1000*MEGmeantime5HzCON[3,],df=20)
lines(spl5cCON$x,spl5cCON$y,col='blue',lwd=2,lty=3)
spl5dCON <- smooth.spline(times,1000*MEGmeantime5HzCON[4,],df=20)
lines(spl5dCON$x,spl5dCON$y,col='orange',lwd=2,lty=3)

plotlims <- c(-1,8,0,2)  # define the x and y limits of the plot (minx,maxx,miny,maxy)
ticklocsx <- -1:8    # locations of tick marks on x axis
ticklocsy <- seq(0,2,1)    # locations of tick marks on y axis
ticklabelsx <- -1:8        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2) 
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx) 
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (s)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    # titles for axes
title(ylab="Masking ratio", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

lines(spl5b$x,spl5a$y/spl5c$y,col='blue',lwd=2)
lines(spl5b$x,spl5a$y/spl5d$y,col='orange',lwd=2)

lines(spl5bASD$x,spl5aASD$y/spl5cASD$y,col='blue',lwd=2,lty=2)
lines(spl5bASD$x,spl5aASD$y/spl5dASD$y,col='orange',lwd=2,lty=2)

lines(spl5bCON$x,spl5aCON$y/spl5cCON$y,col='blue',lwd=2,lty=3)
lines(spl5bCON$x,spl5aCON$y/spl5dCON$y,col='orange',lwd=2,lty=3)

plotlims <- c(-1,8,0,30)  # define the x and y limits of the plot (minx,maxx,miny,maxy)
ticklocsx <- -1:8    # locations of tick marks on x axis
ticklocsy <- seq(0,30,10)    # locations of tick marks on y axis
ticklabelsx <- -1:8        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2) 
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx) 
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (s)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

# lines(times,1000*MEGmeantime7Hz[1,],lwd=2,col='red')
# lines(times,1000*MEGmeantime7Hz[2,],lwd=2,col='green')
# lines(times,1000*MEGmeantime7Hz[3,],lwd=2,col='blue')
# lines(times,1000*MEGmeantime7Hz[4,],lwd=2,col='orange')
spl7a <- smooth.spline(times,1000*MEGmeantime7Hz[1,],df=20)
lines(spl7a$x,spl7a$y,col='red',lwd=2)
spl7b <- smooth.spline(times,1000*MEGmeantime7Hz[2,],df=20)
lines(spl7b$x,spl7b$y,col='green',lwd=2)
spl7c <- smooth.spline(times,1000*MEGmeantime7Hz[3,],df=20)
lines(spl7c$x,spl7c$y,col='blue',lwd=2)
spl7d <- smooth.spline(times,1000*MEGmeantime7Hz[4,],df=20)
lines(spl7d$x,spl7d$y,col='orange',lwd=2)

# lines(times,1000*MEGmeantime7HzASD[1,],lwd=2,col='red',lty=2)
# lines(times,1000*MEGmeantime7HzASD[2,],lwd=2,col='green',lty=2)
# lines(times,1000*MEGmeantime7HzASD[3,],lwd=2,col='blue',lty=2)
# lines(times,1000*MEGmeantime7HzASD[4,],lwd=2,col='orange',lty=2)
spl7aASD <- smooth.spline(times,1000*MEGmeantime7HzASD[1,],df=20)
lines(spl7aASD$x,spl7aASD$y,col='red',lwd=2,lty=2)
spl7bASD <- smooth.spline(times,1000*MEGmeantime7HzASD[2,],df=20)
lines(spl7bASD$x,spl7bASD$y,col='green',lwd=2,lty=2)
spl7cASD <- smooth.spline(times,1000*MEGmeantime7HzASD[3,],df=20)
lines(spl7cASD$x,spl7cASD$y,col='blue',lwd=2,lty=2)
spl7dASD <- smooth.spline(times,1000*MEGmeantime7HzASD[4,],df=20)
lines(spl7dASD$x,spl7dASD$y,col='orange',lwd=2,lty=2)

# lines(times,1000*MEGmeantime7HzCON[1,],lwd=2,col='red',lty=3)
# lines(times,1000*MEGmeantime7HzCON[2,],lwd=2,col='green',lty=3)
# lines(times,1000*MEGmeantime7HzCON[3,],lwd=2,col='blue',lty=3)
# lines(times,1000*MEGmeantime7HzCON[4,],lwd=2,col='orange',lty=3)
spl7aCON <- smooth.spline(times,1000*MEGmeantime7HzCON[1,],df=20)
lines(spl7aCON$x,spl7aCON$y,col='red',lwd=2,lty=3)
spl7bCON <- smooth.spline(times,1000*MEGmeantime7HzCON[2,],df=20)
lines(spl7bCON$x,spl7bCON$y,col='green',lwd=2,lty=3)
spl7cCON <- smooth.spline(times,1000*MEGmeantime7HzCON[3,],df=20)
lines(spl7cCON$x,spl7cCON$y,col='blue',lwd=2,lty=3)
spl7dCON <- smooth.spline(times,1000*MEGmeantime7HzCON[4,],df=20)
lines(spl7dCON$x,spl7dCON$y,col='orange',lwd=2,lty=3)

plotlims <- c(-1,8,0,2)  # define the x and y limits of the plot (minx,maxx,miny,maxy)
ticklocsx <- -1:8    # locations of tick marks on x axis
ticklocsy <- seq(0,2,1)    # locations of tick marks on y axis
ticklabelsx <- -1:8        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2) 
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx) 
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (s)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    # titles for axes
title(ylab="Masking ratio", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

lines(spl7b$x,spl7b$y/spl7c$y,col='blue',lwd=2)
lines(spl7b$x,spl7b$y/spl7d$y,col='orange',lwd=2)

lines(spl7bASD$x,spl7bASD$y/spl7cASD$y,col='blue',lwd=2,lty=2)
lines(spl7bASD$x,spl7bASD$y/spl7dASD$y,col='orange',lwd=2,lty=2)

lines(spl7bCON$x,spl7bCON$y/spl7cCON$y,col='blue',lwd=2,lty=3)
lines(spl7bCON$x,spl7bCON$y/spl7dCON$y,col='orange',lwd=2,lty=3)

# lines(times,1000*meanspec[1,2,],lwd=2,lty=2,col='red')
# lines(times,1000*meanspec[2,2,],lwd=2,lty=2,col='green')
# lines(times,1000*meanspec[3,2,],lwd=2,lty=2,col='blue')
# lines(times,1000*meanspec[4,2,],lwd=2,lty=2,col='orange')

# text(-0.5,3.8,paste0('N=',length(allRnos)),cex=2)
# condlist <- c('5Hz target only','5Hz mask only','5Hz mon','5Hz dich','7Hz target only','7Hz mask only','7Hz mon','7Hz dich')
# if (outputplot>0){legend(0,4,condlist,col=c('red','green','blue','orange'),lty=c(1,1,1,1,2,2,2,2),lwd=2,box.lwd=2)}
